Optimal updating factor in Wang-Landau and metadynamics simulations

### 2016-04-19 doc/invt.tex
even sampling of the optimal schedule

Consider the problem of computing
the distribution, $p^*_i$,
along a discrete quantity, $i$,
for a given system.
%
%The discreteness of $i$ is natural
%in discrete models.
%
For example, $i$ can be the energy, $E$,
in a lattice spin model; or the temperature index
in a simulated tempering simulation\cite{
marinari1992, lyubartsev1992}.
%
For a continuous quantity, $z$,
which is typical in a molecular system,
we can discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin,
$(z, z + dz)$.
%
In both cases,
the distribution is normalized as
$\sum_{i = 1}^n p^*_i = 1$.

@@

Initially,
we set the bias potential to zero,
$v_i \equiv 0$,
%
and equilibrate the system under
a constant updating magnitude,
$\alpha_0 = 10^{-4}$.
%
After this,
we reset the origin of the time, $t$,
and started the target schedule.
%specified by either
%Eq. \eqref{eq:q_opt}
%or
%Eq. \eqref{eq:alpha_invtlambda}.

For the Gaussian updating scheme,
the kernel is truncated at
$\min\{10 \, \sigma, n - 1\}$
and numerically stabilized
using the technique
in Sec. \ref{sec:stabilize_wband}.

Since the error depends on
the underlying sampling process,
we have devised two sampling processes,
a global one and a local one.
%
Both cases were based on
Metropolis algorithm\cite{metropolis1953, newman, frenkel,
landau_binder}:
%
in each step, a new bin index $j$ was proposed
and then accepted with probability
%
$
A(i \to j) = \min\{ 1, \exp(v_i - v_j) \}.
$
However,
in the first and global sampling process,
we choose the destination $j$
randomly out of the $n$ bins.
%
This sampling process
emulates the perfect sampling process
discussed in Sec. \ref{sec:Gamma}
in the long time limit.
%
In the second or local process,
we limit the destination $j$
to the two neighbors of $i$,
$i \pm 1$, with equal probability $1/2$,
to mimic the one-step sampling process
in Sec. \ref{sec:Gamma}.
%

@@

We first studied the inverse-time schedule,
%
\begin{equation}
\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
\label{eq:alpha_invtlambda}
\end{equation}
%
where $\lambda$ is a free parameter,
and $t_0 = 2/(\lambda \, \alpha_0)$.
%
The error of this schedule
can be computed analytically
(cf. Appendix \ref{sec:invt_schedule}),
and thus be tested against the theory.

@@

For the single-bin updating scheme,
we chose $\alpha_0 = 0.0001$
and ran the simulation for $T = 10^8$ steps.
%
As shown in Fig. \ref{fig:errsbin},
the numerical results of the terminal error agreed well with
the analytical prediction by Eq. \eqref{eq:error_invt}.
%
For both global and local sampling processes,
the optimal $\lambda$ occurred around $1.0$,
verifying Eq. \eqref{eq:alpha_invt1}.
%
%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
%represents the relative updating magnitude.
%%
%Figure \ref{fig:errsbin} also shows
%an asymmetry of the error curve.
%A smaller than optimal value of
%the updating magnitude, $1/\lambda$,
%tends to increase the error more rapidly
%than a larger than optimal value,
%suggesting that the over-updating
%is generally preferable to the under-updating
%when the optimal value is unknown.
%
The two error functions
differed only by a multiplicative constant,
again agreeing with the theory.
%
With identical eigenvalues,
$\lambda_1 = \cdots = \lambda_n = 1$,
the error in Eq. \eqref{eq:error_invt}
depends on the sampling process only through
the sum $\sum_{ k = 2 }^n \Gamma_k$,
which gives rise to the above multiplicative constant.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
  }
  \caption{
    \label{fig:errsbin}
    Error, $E$, versus the
    relative updating magnitude,
    %proportionality constant,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the single-bin (WL) scheme.
    %
    The results have been averaged over $1000$ independent runs.
    %
    \note{The figure was produced by \texttt{doc/fig/errsbin.gp}.
      The data were produced by the script \texttt{data/invt2run.py},
      and saved under \texttt{data/singlebin}
      (see the \texttt{Makefile} there).
    }%
  }
\end{center}
\end{figure}

@@

As an example of the multiple-bin schemes,
we chose the nearest-neighbor updating scheme
described in Sec. \ref{sec:nnscheme}.
%
We again find good agreement
between simulation and analytical results.
%
A new feature is that the optimal $\lambda$
depends on the simulation length, $T$.
%
As shown in Fig. \ref{fig:errnnbr},
the normalized error, $(T + t_0) \, \Err$,
hence the optimal value of $1/\lambda$,
depended non-trivially on the simulation length, $T$,
in the global sampling case.
%
%The optimal $1/\lambda$ gradually increased
%with simulation length,
%as discussed in Sec. \ref{sec:optlambda}.
%
For the local sampling case,
the $T$-dependence was hardly visible
because of the predominantly large $\Gamma_k$'s
of the small $k$ modes.
%
The optimal $\lambda$ remained around $1.0$,
as suggested by Eq. \eqref{eq:lambda_nn_onestep}.

@@

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errnnbr.pdf}
  }
  \caption{
    \label{fig:errnnbr}
    Normalized error, $(T + t_0) \, E$,
    versus the %proportionality constant,
    relative updating magnitude,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
    %
    The results have been averaged over $1000$ independent runs.
    %
    \note{This figure was produced by \texttt{doc/fig/errnnbr.gp}.
      The data were produced by \texttt{data/invt2run.py},
      saved under \texttt{data/nb0.24}
      (see the \texttt{Makefile} there).
    }%
  }
\end{center}
\end{figure}

%\subsection{Gaussian (metadynamics) updating scheme}
%
%
%The error of Gaussian updating scheme is similar to
%that of the nearest neighbor scheme,
%as shown in Fig \ref{fig:err_sig5}.
%%
%The dependency of the simulation length is more prominent.
%%
%Even for local sampling,
%there is now a visible shift of the optimal $\lambda$.
%
%
%\begin{figure}[h]
%\begin{center}
%  \makebox[\linewidth][c]{
%    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
%  }
%  \caption{
%    \label{fig:err_sig5}
%    Normalized error, $(T + t_0) \, E$,
%    versus the %proportionality constant,
%    relative updating magnitude,
%    $1/\lambda$,
%    in Eq. \eqref{eq:alpha_invtlambda}
%    for the Gaussian updating scheme with $\sigma = 5$ bins.
%    %
%    The results have been averaged over $1000$ independent runs.
%  }
%\end{center}
%\end{figure}
%

@@

We now turn to the optimal schedule
predicted by Eq. \eqref{eq:q_opt}.
%
For the Gaussian updating scheme
with $\sigma = 10$,
we computed the schedule,
and compared it with the inverse-time schedule
Eq. \eqref{eq:alpha_invtlambda}
with the optimal $\lambda$.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/optacmp.pdf}
  }
  \caption{
    \label{fig:optacmp}
    Optimal schedules from Eq. \eqref{eq:q_opt},
    versus the optimized inverse-time schedule
    Eq. \eqref{eq:alpha_invtlambda}
    for the Gaussian updating scheme
    with $\sigma = 10$.
    %
    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
      The data were computed by the program \texttt{prog/predict},
      see particularly the function \texttt{intq\_save()}
      in \texttt{prog/intq.h}.
      The data were saved under the directory \texttt{data/opta}.
      See the \texttt{make} object \texttt{sigsingle}
      in the \texttt{Makefile} there.
    }%
  }
\end{center}
\end{figure}

As shown in Fig. \ref{fig:optacmp}(a),
the optimal schedule is a more complex curve
than the optimized inverse-time schedule.
%
In Fig. \ref{fig:optacmp}(b),
we show that in the former case,
the rate of change of $\alpha^{-1}(t)$
increases gradually
from $\lambda(0)$ to $\lambda(T) \approx 1$,
as discussed in Sec. \ref{sec:optschedule}.
%
The errors in the two cases were
$2.7 \times 10^{-7}$
and
$6 \times 10^{-7}$,
respectively,
for the global sampling process,
%
or
$9.1 \times 10^{-5}$
and
$1.7 \times 10^{-4}$,
respectively
for the local sampling process,
showing the superiority of
the optimal schedule from Eq. \eqref{eq:q_opt}.

On the other hand,
the optimal schedules of the bandpass updating schemes
are always given by the inverse-time formula,
Eq. \eqref{eq:alpha_invt1},
with the optimal $t_0$
given by Eq. \eqref{eq:t0_sinc},
as shown in Fig. \ref{fig:sinc}(a).

In Fig. \ref{fig:sinc}(b),
we show that the updating kernels
of the bandpass updating schemes
are similar to Gaussian ones
except the oscillatory tails.
%
The equivalent width, $\sigma$, is roughly given
\begin{equation}
  \sigma
  \approx
  \frac
  {
    n
  }
  {
    \sqrt{ 2 \, \pi } \, g \, K
  }
  ,
\label{eq:sigma_equiv}
\end{equation}
where $g = 2$ for the periodic case
or $g = 1$ otherwise.
%
The kernel is a zigzag in the non-periodic case
because of the term $(-1)^{n+K+l-1}$
in Eq. \eqref{eq:mu_sinc_refl}.
%
However, as shown in the inset of Fig. \ref{fig:sinc}(b),
the ruggedness disappears
in the actual updating matrix, Eq. \eqref{eq:w_band_refl},
as the above oscillatory term for $l = i - j$
is canceled by either the term for $l = i + j - 1$
or that for $l = i + j - 2 \, n - 1$,

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
  }
  \caption{
    \label{fig:sinc}
    Bandpass updating schemes.
    (a) Optimal schedule for a periodic variable
    with $K = 3$,
    and that for a non-periodic variable
    with $K = 6$.
    (b) Updating kernels
    compared to a Gaussian with a similar width,
    $\sigma = 5$.
    The lines are a guide to the eyes.
    Inset. Updating matrix in the non-periodic case
    is free of the zigzaggedness shown in the kernel.
    %
    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
      The data were computed by the program \texttt{prog/predict},
      see particularly the function \texttt{intq\_save()}
      in \texttt{prog/intq.h} for panel (a),
      and the functions \texttt{savewin()}
      and \texttt{savewinmat()}
      in \texttt{prog/invt.h} for panel (b),
      The data were saved under the directory \texttt{data/sinc}.
      See the \texttt{make} object \texttt{sincsingle}
      in the \texttt{Makefile} there.
    }%
  }
\end{center}
\end{figure}

@@

For a finite-length simulation,
the optimal updating scheme
depends on the simulation length, $T$,
and other parameters.
%
To illustrate this,
we compare Gaussian updating schemes
with different width, $\sigma$.
%
For the global sampling process,
we see from Fig. \ref{fig:errsigscan}(a)
that the minimal error was achieved
by the widest window for the shorter simulation
of $T = 10^8$.
%
However, as the simulation lengthens,
the minimum at $\sigma = 0$ became more favorable.
%
However,
the shift was hard to detect
for the local sampling scheme,
as shown in Fig. \ref{fig:errsigscan}(b).

In contrast,
the normalized error of the bandpass updating scheme
was insensitive to the simulation length,
in agreement with Eq. \eqref{eq:error_sinc}.
%
Further, in all cases the bandpass updating scheme
appeared to be superior to the Gaussian one.

@@

%\subsection{\label{sec:measure_gamma}
%Measurement of the integrals
%of the autocorrelation functions}

%To use Eq. \eqref{eq:q_opt} in a real simulation,
%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
%
While the updating scheme alone
determines the eigenmodes and $\lambda_k$'s,
the values of $\Gamma_k$'s depend also on the system
and the underlying sampling process (MC or MD).
%
%Below is a technique of measuring $\Gamma_k$'s.
%
%The idea is to run
To measure $\Gamma_k$'s, we can run
a simulation with a constant schedule,
$\alpha(t) = \alpha_0$,
and then use Eq. \eqref{eq:y2_eql}.
%
%A practical difficulty is that
However,
we often do not know the intrinsic distribution, $p^*_i$,
in advance to compute
the shifted bias potential, $v_i$,
from the original one, $V_i$
(cf. Sec. \ref{sec:FDS}).
%
To circumvent the problem, we
modify Eq. \eqref{eq:y2_eql} as
%
\begin{equation}
  \operatorname{var} Y_k
  =
  \frac{1}{2}
  \alpha \, \Gamma_k \, \lambda_k,
\label{eq:varY}
\end{equation}
%
where
``$\operatorname{var}$''
means the variance from time averaging,
and
%
\begin{align}
Y_k
&=
\sum_{ i = 1 }^n
  \phi_{k i} \, (V_i - \bar V),
\label{eq:Y_def}
\\
\bar V
&=
\sum_{ j = 1 }^n p_j \, V_j.
\notag
\end{align}

To show Eq. \eqref{eq:varY},
we first use Eq. \eqref{eq:v_def}
to rewrite Eq. \eqref{eq:x_def}
as
$$
\begin{aligned}
  x_i
  &=
  (V_i - \bar V) + s_i,
  \\
  s_i
  &=
  \ln \frac{ p_i } { p^*_i }
  -
  \sum_{ j = 1 }^n
    p_j \, \ln \frac{ p_j } { p^*_j }.
\end{aligned}
$$
Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
we have
$$
y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
$$
Since the second term on the right-hand side
is a constant, and the average of $y_k$ is zero,
we have
$$
\operatorname{var} Y_k
=
\operatorname{var} y_k
=
\langle y_k^2 \rangle.
$$
Finally, by using Eq. \eqref{eq:y2_eql},
we reach Eq. \eqref{eq:varY}.

@@

%\subsection{\label{sec:program_list}
%Program
%}

@@

%A practical problem of using homogeneous updating scheme
%is the following.
%
A homogeneous updating scheme
specified by an arbitrary kernel
is not necessarily stable,
i.e., some eigenvalues may be negative.
%
%
\note{For example,
  %the nearest-neighbor updating scheme
  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
  %
  %Even
  the Gaussian updating scheme can be unstable
  if the updating kernel is truncated.
}%
%
Here we show how to minimally modify
the updating kernel
to stabilize it.
%
\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
}

@@

For the nearest-neighbor updating scheme
introduced in Sec. \ref{sec:nnscheme},
we can find an analytical expression
of the optimal value of $\lambda$
in the limit of large $n$.
%
In this case,
we convert Eq.
\eqref{eq:error_asym_invt}
%and
%\eqref{eq:optimal_lambda_approx}
to an integral,
$\sum_i \to \frac{2 \, n}{\pi} \int dp$, and
%
\begin{align}
\Err_A
=
\frac{2 \, n}{\pi \, T}
\int_0^{\pi/2}
  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\, dp.
\notag
%\label{eq:error_nn_asym_int}
\end{align}

>>>>

Consider the problem of computing
the distribution, $p^*_i$,
along a discrete quantity, $i$,
for a given system.
%
%The discreteness of $i$ is natural
%in discrete models.
%
For example, $i$ can be the energy, $E$,
in a lattice spin model; or the temperature index
in a simulated tempering simulation\cite{
marinari1992, lyubartsev1992}.
%
For a continuous quantity, $z$,
which is typical in a molecular system,
we can discretize $z$
such that the integer $i$ represents
the index of a small interval, or a bin,
$(z, z + \delta z)$.
%
In both cases,
the distribution is normalized as
$\sum_{i = 1}^n p^*_i = 1$.

@@

Initially,
we set the bias potential to zero,
$V_i \equiv 0$,
%
and equilibrate the system under
a constant updating magnitude,
$\alpha_0 = 10^{-4}$.
%
After this,
we reset the origin of the time, $t$,
and started the testing schedule.
%specified by either
%Eq. \eqref{eq:q_opt}
%or
%Eq. \eqref{eq:alpha_invtlambda}.

For the Gaussian updating scheme,
the kernel was truncated at
$b = \min\{10 \, \sigma, n - 1\}$
and numerically stabilized
using the technique
in Appendix \ref{sec:stabilize_wband}.

Since the error depends on
the underlying sampling process,
we tested two sampling processes,
a global one and a local one.
%
Both are based on
Metropolis MC algorithm\cite{
  metropolis1953, newman, frenkel,
  landau_binder}:
%
in each step, a new bin index $j$ is proposed
and then accepted with probability
%
$
A(i \to j) = \min\{ 1, \exp(V_i - V_j) \}.
$
However,
in the first or global sampling process,
we choose the destination $j$
uniformly out of the $n$ bins.
%
This process
emulates the perfect sampling process
in Sec. \ref{sec:Gamma} asymptotically.
%
In the second or local sampling process,
$j$ is either $i - 1$ or $i + 1$
(adjusted by the boundary condition)
with equal probability $1/2$,
which mimics the one-step sampling process
in Sec. \ref{sec:Gamma}.
%

@@

We first studied the inverse-time schedule,
%
\begin{equation}
\alpha(t) = \frac{1}{\lambda \, (t + t_0) },
\label{eq:alpha_invtlambda}
\end{equation}
%
where $\lambda$ is a free parameter,
and $t_0 = 2/(\lambda \, \alpha_0)$.
%
The error of this schedule
can be computed analytically
(cf. Appendix \ref{sec:invt_schedule}),
and tested against simulation.

@@

For the single-bin updating scheme,
we chose $\alpha_0 = 10^{-4}$
and ran the simulation for $T = 10^8$ steps.
%
As shown in Fig. \ref{fig:errsbin},
the numerical results of the terminal error agreed well with
the analytical prediction by Eq. \eqref{eq:error_invt}.
%
For both global and local sampling processes,
the optimal $\lambda$ occurred around $1.0$,
verifying the optimality of Eq. \eqref{eq:alpha_invt1}.
%
%Recall that $1/\lambda$ in Eq. \eqref{eq:alpha_invtlambda}
%represents the relative updating magnitude.
%%
%Figure \ref{fig:errsbin} also shows
%an asymmetry of the error curve.
%A smaller than optimal value of
%the updating magnitude, $1/\lambda$,
%tends to increase the error more rapidly
%than a larger than optimal value,
%suggesting that the over-updating
%is generally preferable to the under-updating
%when the optimal value is unknown.
%
The two error functions
differed only by a multiplicative constant,
which again agreed with the theory:
%
with identical eigenvalues,
$\lambda_2 = \cdots = \lambda_n = 1$,
the error in Eq. \eqref{eq:error_invt}
depends on the sampling process only through
the sum $\sum_{ k = 2 }^n \Gamma_k$,
which gives rise to the above multiplicative constant.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsbin.pdf}
  }
  \caption{
    \label{fig:errsbin}
    Error, $E$, versus the
    relative updating magnitude,
    %proportionality constant,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the single-bin (WL) scheme.
    %
    The results were averaged over $1000$ independent runs.
    %
    \note{The figure was produced by \texttt{doc/fig/errsbin.gp}.
      The data were produced by the script \texttt{data/invt2run.py},
      and saved under \texttt{data/singlebin}
      (see the \texttt{Makefile} there).
    }%
  }
\end{center}
\end{figure}

@@

As an example of the multiple-bin schemes,
we chose the nearest-neighbor updating scheme
described in Sec. \ref{sec:nnscheme}.
%
We choose $\mu_1 = 0.24$ to the smallest eigenvalue
close to zero.
%
Again, simulation and theory agreed well,
as shown in Fig. \ref{fig:errnnbr}.
%
The normalized error
and the optimal value of $1/\lambda$,
depended on the simulation length, $T$,
in the global sampling case,
because of the near zero eigenvalues,
as discussed in Appendix \ref{sec:invt_schedule}.
%
%The optimal $1/\lambda$ gradually increased
%with simulation length,
%as discussed in Sec. \ref{sec:optlambda}.
%
However, for the local sampling case,
the $T$-dependence was suppressed
because of the dominance of large $\Gamma_k$'s
of the small-$k$ modes.
%
The optimal $\lambda$ remained around $1.0$,
as suggested by Eq. \eqref{eq:lambda_nn_onestep}.

@@

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/errnnbr.pdf}
  }
  \caption{
    \label{fig:errnnbr}
    Normalized error, $(T + t_0) \, E$,
    versus the %proportionality constant,
    relative updating magnitude,
    $1/\lambda$,
    in Eq. \eqref{eq:alpha_invtlambda}
    for the nearest-neighbor updating scheme with $\mu_1 = 0.24$.
    %
    The results were averaged over $1000$ independent runs.
    %
    \note{This figure was produced by \texttt{doc/fig/errnnbr.gp}.
      The data were produced by \texttt{data/invt2run.py},
      saved under \texttt{data/nb0.24}
      (see the \texttt{Makefile} there).
    }%
  }
\end{center}
\end{figure}

%\subsection{Gaussian (metadynamics) updating scheme}
%
%
%The error of Gaussian updating scheme is similar to
%that of the nearest neighbor scheme,
%as shown in Fig \ref{fig:err_sig5}.
%%
%The dependency of the simulation length is more prominent.
%%
%Even for local sampling,
%there is now a visible shift of the optimal $\lambda$.
%
%
%\begin{figure}[h]
%\begin{center}
%  \makebox[\linewidth][c]{
%    \includegraphics[angle=0, width=0.95\linewidth]{fig/errsig5.pdf}
%  }
%  \caption{
%    \label{fig:err_sig5}
%    Normalized error, $(T + t_0) \, E$,
%    versus the %proportionality constant,
%    relative updating magnitude,
%    $1/\lambda$,
%    in Eq. \eqref{eq:alpha_invtlambda}
%    for the Gaussian updating scheme with $\sigma = 5$ bins.
%    %
%    The results were averaged over $1000$ independent runs.
%  }
%\end{center}
%\end{figure}
%

@@

We now turn to the optimal schedule
predicted by Eq. \eqref{eq:q_opt}.
%
For the Gaussian updating scheme
with $\sigma = 10$,
we computed the schedule for the local sampling process,
and compared it with the inverse-time schedule
Eq. \eqref{eq:alpha_invtlambda}
with the optimal $\lambda$.
%
As shown in Fig. \ref{fig:optacmp}(a),
the optimal schedule is a more complex curve
than the optimized inverse-time schedule.
%
The complexity of the optimal schedule
corresponds to the non-exponential tail
of the mass distribution function,
as shown Fig. \ref{fig:optacmp}(b)
[cf. Sec. \ref{sec:mass_distr}].
%
The errors in the two cases were
$9.1 \times 10^{-5}$
and
$1.7 \times 10^{-4}$,
respectively,
%for the local sampling process,
showing the superiority of
the optimal schedule from Eq. \eqref{eq:q_opt}.
%
Besides, we verified the scaling property
discussed in Sec. \ref{sec:mass_distr},
by comparison with a simulation [Fig. \ref{fig:optacmp}(a)]
of half length, $T' = 5 \times 10^7$
and doubled initial updating magnitude,
$\alpha_0 = 2 \times 10^{-4}$,
which left the normalized mass distribution invariant,
as shown in Fig. \ref{fig:optacmp}(b).

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/optacmp.pdf}
  }
  \caption{
    \label{fig:optacmp}
    Optimal schedules from Eq. \eqref{eq:q_opt},
    versus the optimized inverse-time schedule
    Eq. \eqref{eq:alpha_invtlambda}
    for the Gaussian updating scheme
    with $\sigma = 10$
    and the local updating process.
    %
    \note{This figure was produced by \texttt{doc/fig/optacmp.gp}.
      The data were computed by the program \texttt{prog/predict},
      see particularly the function \texttt{intq\_save()}
      in \texttt{prog/intq.h}.
      The data were saved under the directory \texttt{data/opta}.
      See the \texttt{make} object \texttt{sigsingle}
      in the \texttt{Makefile} there.
    }%
  }
\end{center}
\end{figure}

On the other hand,
the optimal schedules of the bandpass updating schemes
are always given by the inverse-time formula,
Eq. \eqref{eq:alpha_invt1},
with $t_0$ given by Eq. \eqref{eq:t0_sinc},
as shown in Fig. \ref{fig:sinc}(a).
%
In Fig. \ref{fig:sinc}(b),
we show that the updating kernels
of the bandpass updating schemes
are similar to Gaussian ones
except the oscillatory tails.
%
The equivalent width, $\sigma$, is roughly
%
\begin{equation}
  \sigma
  \approx
  \frac
  {
    n
  }
  {
    \sqrt{ 2 \, \pi } \, g \, K
  }
  ,
\label{eq:sigma_equiv}
\end{equation}
%
where $g = 2$ for the periodic case
or $g = 1$ otherwise.
%
The kernel is a zigzag in the non-periodic case
because of the term $(-1)^{n+K+l-1}$
in Eq. \eqref{eq:mu_sinc_refl}.
%
However, as shown in the inset, % of Fig. \ref{fig:sinc}(b),
the ruggedness disappears
in the updating matrix by Eq. \eqref{eq:w_band_refl},
as the above oscillatory term for $l = i - j$
is canceled by the term for either $l = i + j - 1$
or $l = i + j - 2 \, n - 1$
with the opposite sign.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/sinc.pdf}
  }
  \caption{
    \label{fig:sinc}
    Bandpass updating schemes.
    (a) Optimal schedule for a periodic variable
    with $K = 3$,
    and that for a non-periodic variable
    with $K = 6$.
    (b) Updating kernels
    compared to a Gaussian with a similar width,
    $\sigma = 6$.
    %
    The local sampling process was used.
    The lines are a guide to the eyes.
    Inset. Updating matrix in the non-periodic case
    is free of the zigzaggedness shown in the kernel.
    %
    \note{This figure was produced by \texttt{doc/fig/sinc.gp}.
      The data were computed by the program \texttt{prog/predict},
      see particularly the function \texttt{intq\_save()}
      in \texttt{prog/intq.h} for panel (a),
      and the functions \texttt{savewin()}
      and \texttt{savewinmat()}
      in \texttt{prog/invt.h} for panel (b),
      The data were saved under the directory \texttt{data/sinc}.
      See the \texttt{make} object \texttt{sincsingle}
      in the \texttt{Makefile} there.
    }%
  }
\end{center}
\end{figure}

@@

For a finite-length simulation,
the optimal updating scheme
depends on the simulation length, $T$,
and other parameters.
%
As an example,
we compare Gaussian updating schemes
with different widths, $\sigma$.
%
As shown in Fig. \ref{fig:errsigscan}(a)
that the minimal error was achieved
by the widest window for the shorter simulation
of $T = 10^8$,
for the global sampling process.
%
As the simulation lengthens,
the minimum at $\sigma = 0$ became more favorable.
%
However,
the minimum stayed at zero width,
as shown in Fig. \ref{fig:errsigscan}(b).

In contrast,
the normalized error of the bandpass updating schemes
were insensitive to the simulation length,
in agreement with Eq. \eqref{eq:error_sinc}.
%
Besides, the bandpass updating schemes
yielded smaller errors than
the corresponding Gaussian ones.

@@

\subsection{\label{sec:measure_gamma}
Measurement of the integrals
of the autocorrelation functions}

%To use Eq. \eqref{eq:q_opt} in a real simulation,
%we need to know both $\lambda_k$'s and $\Gamma_k$'s.
%
While the updating scheme alone
determines the eigenmodes and $\lambda_k$'s,
the values of $\Gamma_k$'s depend also on the system
and the underlying sampling process (MC or MD).
%
%Below is a technique of measuring $\Gamma_k$'s.
%
%The idea is to run
To measure $\Gamma_k$'s, we can run
a simulation with a constant schedule,
$\alpha(t) = \alpha_0$,
and then use Eq. \eqref{eq:y2_eql}.
%
%A practical difficulty is that
However,
we often do not know the intrinsic distribution, $p^*_i$,
in advance to compute
the shifted bias potential, $v_i$,
from the original one, $V_i$
(cf. Sec. \ref{sec:FDS}).
%
To circumvent the problem, we
modify Eq. \eqref{eq:y2_eql} as
%
\begin{equation}
  \operatorname{var} Y_k
  =
  \frac{1}{2}
  \alpha \, \Gamma_k \, \lambda_k,
\label{eq:varY}
\end{equation}
%
where
``$\operatorname{var}$''
means the variance from time averaging,
and
%
\begin{align}
Y_k
&=
\sum_{ i = 1 }^n
  \phi_{k i} \, (V_i - \bar V),
\label{eq:Y_def}
\end{align}
%
with
$
\bar V
\equiv
\sum_{ j = 1 }^n p_j \, V_j.
$

To show Eq. \eqref{eq:varY},
we first use Eq. \eqref{eq:v_def}
to rewrite Eq. \eqref{eq:x_def}
as
$
  x_i
  =
  (V_i - \bar V) + s_i,
$
where
$
  s_i
  \equiv
  \ln \frac{ p_i } { p^*_i }
  -
  \sum_{ j = 1 }^n
    p_j \, \ln \frac{ p_j } { p^*_j }.
$
Thus, by Eqs. \eqref{eq:y_def} and \eqref{eq:Y_def},
we have
$$
y_k = Y_k + \sum_{i = 1}^n \phi_{k i} \, s_i.
$$
Since the second term on the right-hand side
is a constant, and the average of $y_k$ is zero,
we have
$$
\operatorname{var} Y_k
=
\operatorname{var} y_k
=
\langle y_k^2 \rangle.
$$
Finally, by using Eq. \eqref{eq:y2_eql},
we reach Eq. \eqref{eq:varY}.

@@

As an example, we used the method to compute
the integrals, $\Gamma_k$'s, from an MD simulation,
and compared them to those from
the local MC sampling process
[cf. Sec. \ref{sec:results_system}].
%
As shown in Fig. \ref{fig:gamcmp},
the larger values of $\Gamma_k$'s
with smaller $k$'s
from the two sampling processes
were almost proportional,
suggesting similarity 
of local sampling processes.

\begin{figure}[h]
\begin{center}
  \makebox[\linewidth][c]{
    \includegraphics[angle=0, width=0.95\linewidth]{fig/gamcmp.pdf}
  }
  \caption{
    \label{fig:gamcmp}
    Integrals of the autocorrelation functions
    from an MD simulation versus those
    from the local MC simulation.
    %
    The MD simulation was ran for $10^{10}$ steps with
    the step size $\Delta t_\mathrm{MD} = 0.005$.
    %
    The bin size $\Delta z = 2\, \pi /n$ with $n = 100$,
    and the temperature was $1.0$.
    %
    The velocity rescaling\cite{bussi2007} thermostat was
    applied every MD step with the viscocity set to 20.
    %
    The reflective boundary condition was used.
    %
    \note{The figure was produced by \texttt{doc/fig/gamcmp.gp}.
      Thee data were saved under \texttt{data/gamcmp}
      (see the \texttt{Makefile} there).
    }%
  }
\end{center}
\end{figure}

\subsection{\label{sec:program_list}
Program
}

@@

%A practical problem of using homogeneous updating scheme
%is the following.
%
A homogeneous updating scheme
specified by an arbitrary kernel
is not necessarily stable,
i.e., some eigenvalues may be negative.
%
%
%\note{
  %For example,
  %the nearest-neighbor updating scheme
  %is unstable if Eq. \eqref{eq:nn_stable} is violated.
  %
  Even
  the Gaussian updating scheme can be unstable
  if the updating kernel is truncated.
%}%
%
Here we show how to minimally modify
the updating kernel
to stabilize it.
%
\note{The relative code is \texttt{trimwindow()} in \texttt{invt.h}.
}

@@

For the nearest-neighbor updating scheme
introduced in Sec. \ref{sec:nnscheme},
we can find an analytical expression
of the optimal value of $\lambda$
in the large-$n$ limit.
%
In this case,
we
drop the second term in the square brackets
of Eq. \eqref{eq:error_asym_invt},
which vanishes asymptotically if $2 \, \nu_k > 1$,
and convert the asymptotic error
to an integral:
%$\sum_i \to \frac{2 \, n}{\pi} \int_0^{\pi/2} dp$, and
%
\begin{align}
\Err_A
=
\frac{2 \, n}{\pi \, T}
\int_0^{\pi/2}
  \frac{ \Gamma(p) \, \left(1 - 4 \, \mu_1 \, \sin^2 p \right)^2    }
       {   \lambda \, \left(2 - 8 \, \mu_1 \, \sin^2 p - \lambda \right) }
\, dp.
\notag
%\label{eq:error_nn_asym_int}
\end{align}
